Reproducible Research Project
=============================
This is my firstR markdown file

Loading and preprocessing the data

Download, unzip and load data into data frame data.

```{r}
activity<- read.csv("activity.csv")
```
A brief description of the dataset

```{r}
head(activity)
```

```{r}
dim(activity)
```

```{r}

str(activity)
```

```{r}
summary(activity)
```

What is mean total number of steps taken per day?
=================================================

Sum steps by day, create Histogram, and calculate mean and median.
```{r}
steps_per_day <- aggregate(steps ~ date, activity, sum)
hist(steps_per_day$steps, main = paste("Total Steps Each Day"), col="blue", xlab="Number of Steps")
```

```{r}
rmean <- mean(steps_per_day$steps)
rmean
rmedian <- median(steps_per_day$steps)
rmean
```

What is the average daily activity pattern?
==============================================
Calculate average steps for each interval for all days.
Plot the Average Number Steps per Day by Interval.
Find interval with most average steps.
```{r}
steps_per_interval <- aggregate(steps ~ interval, activity, mean)
plot(steps_per_interval$interval,steps_per_interval$steps, type="l", xlab="Interval", ylab="Number of Steps",main="Average Number of Steps per Day by Interval")
```

```{r}
library(stats)
```

```{r}
max_interval <- steps_per_interval[which.max(steps_per_interval$steps),1]
max_interval
```

Impute missing values.
======================

Compare imputed to non-imputed data.

Missing data needed to be imputed. Only a simple imputation approach was required for this assignment. Missing values were imputed by inserting the average for each interval. Thus, if interval 10 was missing on 10-02-2012, the average for that interval for all days (0.1320755), replaced the NA.

```{r}
incomplete <- sum(!complete.cases(activity))
imputed_data <- transform(activity, steps = ifelse(is.na(activity$steps), steps_per_interval$steps[match(activity$interval, steps_per_interval$interval)], activity$steps))
```
Zeroes were imputed for 10-01-2012 because it was the first day and would have been over 9,000 steps higher than the following day, which had only 126 steps. NAs then were assumed to be zeros to fit the rising trend of the data.

```{r}
imputed_data[as.character(imputed_data$date) == "2012-10-01", 1] <- 0
```

Recount total steps by day and create Histogram.

```{r}
steps_per_day_i <- aggregate(steps ~ date, imputed_data, sum)
hist(steps_per_day_i$steps, main = paste("Total Steps Each Day"), col="green", xlab="Number of Steps")

# Create Histogram to show difference. 
hist(steps_per_day$steps, main = paste("Total Steps Each Day"), col="blue", xlab="Number of Steps", add=T)
legend("topright", c("Imputed", "non-imputed"), col=c("green", "blue"), lwd=10)

```

Calculate new mean and median for imputed data.
```{r}
rmean.i <- mean(steps_per_day_i$steps)
rmean.i
rmedian.i <- median(steps_per_day_i$steps)
rmedian.i
```
Calculate difference between imputed and non-imputed data.

```{r}
mean_diff <- rmean.i - rmean
mean_diff
med_diff <- rmedian.i - rmedian
med_diff
```

Calculate total difference.
```{r}
total_diff <- sum(steps_per_day_i$steps) - sum(steps_per_day$steps)
total_diff
```


Create a plot to compare and contrast number of steps between the week and weekend. There is a higher peak earlier on weekdays, and more overall activity on weekends.
```{r}
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", 
"Friday")
imputed_data$dow = as.factor(ifelse(is.element(weekdays(as.Date(imputed_data$date)),weekdays), "Weekday", "Weekend"))

steps_per_interval_i <- aggregate(steps ~ interval + dow, imputed_data, mean)


library(lattice)

xyplot(steps_per_interval_i$steps ~ steps_per_interval_i$interval|steps_per_interval_i$dow, main="Average Steps per Day by Interval",xlab="Interval", ylab="Steps",layout=c(1,2), type="l")

```


